\documentclass[12pt]{report}

\usepackage{url}
\usepackage{xspace}
\usepackage{fancybox}
\usepackage{cuthesis}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{csquotes}
\usepackage{multirow}
\usepackage{pdflscape}
\usepackage{subfigure}



\author{Everton da S. Maldonado}
\title {Identifying Self-Admitted Technical Debt}
\degree{Master of Applied Science in Software Engineering}
\dept  {Computer Science and Software Engineering}

\newcommand{\SATD}{self-admitted technical debt\xspace}

\newcommand{\chapterIIIrqi}{\textbf{RQ: What are the types of self-admitted technical debt? How frequent are the different types of self-admitted technical debt in the studied projects ?}}

\newcommand{\chapterIVrqi}{\textbf{RQ1. Is it possible to more accurately detect \SATD using NLP techniques?\\}}
\newcommand{\chapterIVrqii}{\textbf{RQ2. What are the most impactful words in the classification of \SATD?\\}}
\newcommand{\chapterIVrqiii}{\textbf{RQ3. How much training data is required to effectively detect \SATD?\\}}

\newcommand{\conclusionbox}[1]{%
    \vspace{6mm}
    \framebox[0.95\textwidth][c]{%
        \parbox[b]{0.92\textwidth}{%
            {\it #1}
        }
    }
    \vspace{6mm}
}


\begin{document}

\begin{abstract}

Technical debt is a metaphor coined to express the trade off between productivity and quality, e.g., when developers take shortcuts or perform quick hacks during the development of software projects. These non optimal solutions are often implemented to allow the project to move faster in the short term, at the cost of increased maintenance in the future. The accumulation of technical debt during the ever changing life-cycle of a project is unavoidable, and if not properly managed can severely hinder the development of the project. To help alleviate the impact of technical debt, a number of studies focused on the detection of technical debt. However, a recent study has shown that one possible source to detect technical debt is using source code comments, also referred to as self-admitted technical debt. Therefore, in this dissertation we use empirical studies and NLP techniques to propose an approach to automatically identify \SATD. 

First, we examine source code comments to determine the different types of technical debt, and we propose four simple filtering heuristics to eliminate comments that are not likely to contain technical debt. Then, we read through more than 33K comments, and we find that self-admitted technical debt can be classified into five main types - design debt, defect debt, documentation debt, requirement debt and test debt. In addition, the  most common type of self-admitted technical debt is design debt, making up between 42\% to 84\% of the classified comments. We also make the dataset used in this study publicly available to support future research in the area. 

Second, we leverage the knowledge obtained in our first study to present an approach to automatically identify design and requirement self-admitted technical debt using Natural Language Processing (NLP). We study 10 open source projects: Ant, ArgoUML, Columba, EMF, Hibernate, JEdit, JFreeChart, Jmeter, JRuby and SQuirrel SQL and find that 1) we are able to effectively identify self-admitted technical debt, significantly outperforming state-of-the-art techniques; 2) that words related to sloppy or mediocre source code are the best indicators of design debt, whereas for requirement debt, words related to enhancing or completing tasks are the best indicators; and 3)  we can achieve 90\% of the best classification performance, using as little as 23\% of the comments for both design and requirement self-admitted technical debt, and 80\% of the best performance, using as little as 9\% and 5\% of the comments for design and requirement self-admitted technical debt, respectively.

\end{abstract}

\begin{acknowledgments}
At the conclusion of a stage, you must look back and take the time to thank and to be grateful for those who have been by our side, because happiness is meaningless without someone to share. First, I am grateful to God who gives me strength, guidance and the opportunity to pursue my masters degree. 

I would like to express my gratitude to my advisor Dr. Emad Shihab for his support and dedication during this last two years. Thank you Emad for believing in my potential, for teaching me how to overcome my limitations and that, through hard work, everything is achievable. I must tell you that it has been quite an enjoyable adventure, and that I learned a lot from you. 

Apart from my advisor, I would also like to thank my thesis examiners, Dr. X and Dr. Y, for taking the time to read my thesis and for their valuable suggestions. Also, I would like to acknowledge Dr. Nikolaos Tsantalis for his help in my research, as well as all other faculty members of the Department of Computer Science and Software Engineering, for providing the necessary guidance.

In addition, a big thank you to all my lab mates and everyone else that I had the opportunity to work with. A special thanks to Moiz Arif, Davood Mazinanian, Ahmad Hassan, Samuel Donadelli, Rabe Abdalkareem and Shahriar Rostami. All of you made this experience so much more enjoyable. 

More than anyone else, I would like to thank my wife Fabiana Maldonado. It was your love, continued support and encouragement that made this degree possible. You are more than I could wish for. Also, during this journey, I was very fortunate to count with so many dear friends. I would like to thank Eduardo Lomonaco, K\^{e}nia Balestra, Rodrigo Mayer, Luciana Schurt, Carlos and Antonia Lorenz for being always there for me. 

Lastly, thank you mom, dad, brothers and grandparents, although far away you are always feeling my heart with your love. I dedicate this thesis for you.

\end{acknowledgments}

\chapter{Introduction}
\label{introduction}
\input{introduction.tex}

\chapter{Literature Review}
\label{literature_review}
\input{literature_review.tex}

\chapter{Analyzing Source Code Comments and Different Types of Self-Admitted Technical Debt}
\label{chapter3}
\input{chapter3.tex}

\chapter{Proposing an Approach to Automatically Identify Self-Admitted Technical Debt}
\label{chapter4}
\input{chapter4.tex}

\chapter{Summary, Contributions and Future Work}
\label{conclusion}
\input{conclusion.tex}


\addcontentsline{toc}{chapter}{Bibliography}
\bibliography{bibliography.bib}  
\bibliographystyle{alpha}
\end{document}
